\chapter*{References}
\markboth{References}{}
\addcontentsline{toc}{chapter}{References}

\begin{enumerate}
\item Krishan Yadav, 2026. \textit{DeepSceneLoc - A Visual-Only Framework for Semantic Location Estimation}. Internal project review manuscript, Department of CSE, Parul University.

\item Aditi Sah, 2026. \textit{DeepSceneLoc: A Research Proposal for Visual-Only Semantic Location Estimation}. Internal project review manuscript, Department of CSE, Parul University.

\item Jensi Paneliya, 2026. \textit{DeepSceneLoc: Visual Scene Understanding for Semantic Location Estimation}. Internal project review manuscript, Department of CSE, Parul University.

\item Anuj Kondawar, 2026. \textit{DeepSceneLoc: Semantic Visual Scene Understanding-Based Image Location Estimation without Geographic Metadata}. Internal project review manuscript, Department of CSE, Parul University.

\item Tan, M. \& Le, Q.V., 2019. EfficientNet: Rethinking Model Scaling for Convolutional Neural Networks. In: \textit{Proc. International Conference on Machine Learning (ICML)}. [Online]. Available: \url{https://arxiv.org/abs/1905.11946}

\item Fei-Fei, L., Karpathy, A. \& Johnson, J., 2017. Convolutional Neural Networks for Visual Recognition. Stanford CS231n Course Notes, Stanford University. [Online]. Available: \url{http://cs231n.stanford.edu/}

\item Dosovitskiy, A., et al., 2021. An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale. In: \textit{Proc. International Conference on Learning Representations (ICLR)}. [Online]. Available: \url{https://arxiv.org/abs/2010.11929}

\item Zhou, B., Lapedriza, A., Khosla, A., Oliva, A. \& Torralba, A., 2017. Places: A 10 Million Image Database for Scene Recognition. \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}. [Online]. Available: \url{https://arxiv.org/abs/1610.02055}

\item Saoud, R. \& Larabi, S., 2018. Visual Geo-Localization from Images. In: \textit{Proc. IEEE International Conference on Image Processing (ICIP)}. [Online]. Available: \url{https://ieeexplore.ieee.org/document/8451659}

\item Hausler, S., Garg, S., Xu, M. \& Milford, M., 2021. Patch-NetVLAD: Multi-Scale Fusion of Locally-Global Descriptors for Place Recognition. In: \textit{Proc. IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}. [Online]. Available: \url{https://arxiv.org/abs/2104.14974}

\item Huang, R., et al., 2021. RepVGG: Making VGG-Style ConvNets Great Again. In: \textit{Proc. IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}. [Online]. Available: \url{https://arxiv.org/abs/2101.03697}

\item Chawla, N.V., Bowyer, K.W., Hall, L.O. \& Kegelmeyer, W.P., 2002. SMOTE: Synthetic Minority Over-sampling Technique. \textit{Journal of Artificial Intelligence Research}. [Online]. Available: \url{https://www.jair.org/index.php/jair/article/view/10302}

\item Chen, C., Wang, B., Lu, C.X., Trigoni, N. \& Markham, A., 2022. Deep Learning for Visual Localization and Mapping: A Survey. \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}. [Online]. Available: \url{https://arxiv.org/abs/2006.12562}

\item Jain, A., Verma, C., Kumar, N., et al., 2020. Image Geo-Site Estimation Using Convolutional Auto-Encoder and Multi-Label Support Vector Machine. \textit{IEEE Access}. [Online]. Available: \url{https://ieeexplore.ieee.org/document/9141093}
\end{enumerate}
